Different environment filters shape the functional composition of canopy and understory guilds in an Indian rainforest

### 2020-04-02 Manuscript.Rmd
20200402 without extreme canopy removal

---
title: Different environment filters shape the functional composition of canopy and understory guilds in an Indian rainforest
date: '`r Sys.Date()`'
output:
  bookdown::word_document2:
    reference_docx: ./template/template.docx
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(raster)
library(parallel)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
library(reshape2)
library(ggfortify)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = T, cache.lazy = F)
cv <- function(x, na.rm = T)
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
n = 999 # Random repetitions
sapply(list.files("R", full.names = T), source)
```

```{r data}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-LTD, -SHA, -Sp)
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
PFT <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata)) %>% 
  dplyr::select(strata, SpCode, LA, WD, LT, LDMC, SLA) %>% 
  reshape2::melt(id.vars = c("strata", "SpCode"), 
                 variable.name = "trait", value.name = "traitvalue")
# Matrix L
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
trees <- genTrees() %>% 
  mutate(community = (XY %over% com)[,1]) %>% 
  filter(!is.na(community)) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = n()) %>% 
  filter(community != "C99")
# Matrix R
load(file.path("draft_save", 'R.Rdata'))
# GWM
GWM <- trees %>% 
  left_join(PFT) %>% 
  filter(!is.na(strata)) %>% 
  group_by(strata, community, trait) %>% 
  summarise(GWM = weighted.mean(traitvalue, Abundance, na.rm = T))
# GWM plot
left_join(GWM, R) %>% 
  reshape2::melt(id.vars = c("strata", "community", "trait", "GWM")) %>% 
  filter(variable != "BA") %>% 
  ggplot(aes(value, GWM, col = strata)) +
  facet_grid(trait ~ variable, scales = "free") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom", axis.title = element_blank())
# nGWM
nGWM <- lapply(1:n, function(n) 
  left_join(trees, 
            group_by(PFT, strata) %>% 
              mutate(SpCode = sample(SpCode)),
            by = "SpCode") %>% 
    filter(!is.na(strata)) %>% 
    group_by(strata, community, trait) %>% 
    summarise(nGWM = weighted.mean(traitvalue, Abundance, na.rm = T)) %>% 
    mutate(NullModel = n)) %>% 
  bind_rows()
# nGWM plot
ggplot(nGWM, aes(nGWM, col = community), fill = NA) +
  geom_density(alpha = .5) +
  facet_wrap(~ strata + trait, scales = "free") +
  scale_color_discrete(guide = "none")
# SES
SES <- GWM %>% 
  left_join(group_by(nGWM, strata, community, trait) %>% 
              summarise(nGWMmean = mean(nGWM, na.rm = T),
                        nGWMsd = sd(nGWM, na.rm = T))) %>% 
  mutate(SES = (GWM - nGWMmean)/nGWMsd) %>% 
  dplyr::select(-GWM, -nGWMmean, -nGWMsd) %>% 
  left_join(R)
# SES plot
reshape2::melt(SES, id.vars = c("strata", "community", "trait", "SES")) %>% 
  filter(variable != "BA") %>% 
  ggplot(aes(value, SES, col = strata)) +
  facet_grid(trait ~ variable, scales = "free") +
  geom_point() +
  geom_hline(yintercept = c(-2, 0, 2), linetype = "dashed") +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom", axis.title = element_blank())
```

@@

```{r pca,  fig.cap='Principal Component Analysis (PCA) of species mean trait values (A) and density distribution of species per guild (C canopy, U understory) along the first two PCA axes (B). Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test. See Table  \\@ref(tab:traits) for trait abbreviations.'}
pca.plot <- autoplot(PFT.pca, 
         data = Q %>% 
           left_join(L %>% 
                       dplyr::select(-com) %>% 
                       summarise_all(sum) %>% 
                       t() %>% 
                       data.frame() %>% 
                       rename(abundance =  ".") %>% 
                       mutate(SpCode = row.names(.))) %>% 
           filter(!is.na(WES)), 
         colour = "strata",
         size = "abundance",
         loadings.label.size = 6,
         loadings.label.colour = 'black',
         loadings = T, loadings.label = T, loadings.colour = 'black',
         loadings.label.vjust = 1.2) +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1)) +
  scale_color_discrete("Guild")
spectrum.boxplots <- pca.plot$data %>%
  select(Comp.1, Comp.2, strata) %>%
  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
  melt(id.vars = "strata", variable.name = "spectrum") %>%
  ggplot(aes(strata, value, fill = strata)) +
  ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
  geom_boxplot() +
  facet_wrap(~spectrum, nrow = 2) +
  scale_fill_discrete("Guild")
cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1))
# ggsave(cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1), labels = c("A", "B")),
#        file = "PCA.png",
#        path = './data/',
#        width = 200, height = 150, unit = 'mm',
#        dpi = 300)
```

@@

```{r GWMboxplot, fig.cap="Guild weighted mean (GWM) of specific leaf area (SLA), wood density (WD), leaf dry matter content (LDMC), leaf area (LA) and leaf thickness (LT) for canopy (C) and understory (U) guilds. Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines). Asterisks represent significant differences between the two guilds (*** p-value<0.001).", fig.height=4}
L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  inner_join(Q, by = "SpCode") %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  group_by(strata, community) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
  select(-community, -LES, -WES) %>%
  # select(-community, -SLA, -WD, -LDMC, -LA, -LT) %>% 
  # select(-community) %>% 
  melt(id.vars = "strata", 
       variable.name = "trait",
       value.name = "GWM") %>% 
  ggplot(aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  facet_wrap(~trait, scales = "free", nrow = 1) +
  # facet_wrap(~trait, scales = "free", nrow = 2) +
  xlab("Guild") + scale_fill_discrete(guide = "none") +
  ggpubr::stat_compare_means(aes(label = paste(substr(..p.signif.., 1,3))))
  # ggsave(file = "boxplot.png",
  #        path = './data/',
  #        width = 200, height = 150, unit = 'mm',
  #        dpi = 300)
```

>>>>

---
title: Different environment filters shape the functional composition of canopy and understory guilds in an Indian rainforest
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false  
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
  bookdown::word_document2:
    reference_docx: ./template/template.docx
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(raster)
library(parallel)
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
library(reshape2)
library(ggfortify)
theme_set(bayesplot::theme_default())
opts_chunk$set(echo = F, message = F, warning = F, 
               fig.height = 8, fig.width = 8,
               cache = F, cache.lazy = F)
cv <- function(x, na.rm = T)
  sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm)
n = 999 # Random repetitions
sapply(list.files("R", full.names = T), source)
```

```{r data}
# Sp level PFT
PFT <- genPFT() %>% 
  mutate(LA = log(LA)) %>% 
  mutate(SLA = log(SLA)) %>% 
  select(-CE) %>% 
  group_by(Tree, Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-Tree) %>% 
  group_by(Sp, SpCode) %>% 
  summarise_all(funs(mean(., na.rm = T))) %>% 
  ungroup() %>% 
  select(-LTD, -SHA, -Sp)
# PCA
PFT.pca <- princomp(~ LA + WD + LT + LDMC + SLA, data = PFT, cor = T)
# Matrix Q
PFT <- PFT %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  left_join(genSpecies(), by = "SpCode") %>% 
  mutate(strata = ifelse(Strata == 3, "U", "C")) %>%
  mutate(strata = ifelse(SpCode == "agin", "U", strata)) %>% 
  dplyr::select(strata, SpCode, LA, WD, LT, LDMC, SLA) %>% 
  reshape2::melt(id.vars = c("strata", "SpCode"), 
                 variable.name = "trait", value.name = "traitvalue")
# Matrix L
com <- quadrats(cs = c(30,30))
XY <- genTrees()[c('x', 'y')]
coordinates(XY) <- ~ x + y
proj4string(XY) <- crs(com)
trees <- genTrees() %>% 
  mutate(community = (XY %over% com)[,1]) %>% 
  filter(!is.na(community)) %>% 
  group_by(community, SpCode) %>% 
  summarise(Abundance = n())
# Matrix R
load(file.path("draft_save", 'R.Rdata'))
# GWM
GWM <- trees %>% 
  left_join(PFT) %>% 
  filter(!is.na(strata)) %>% 
  group_by(strata, community, trait) %>% 
  summarise(GWM = weighted.mean(traitvalue, Abundance, na.rm = T))
# GWM plot
left_join(GWM, R) %>% 
  reshape2::melt(id.vars = c("strata", "community", "trait", "GWM")) %>% 
  filter(variable != "BA") %>% 
  ggplot(aes(value, GWM, col = strata)) +
  facet_grid(trait ~ variable, scales = "free") +
  geom_point() +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom", axis.title = element_blank())
# nGWM
nGWM <- lapply(1:n, function(n) 
  left_join(trees, 
            group_by(PFT, trait, strata) %>% 
              mutate(SpCode = sample(SpCode)),
            by = "SpCode") %>% 
    filter(!is.na(strata)) %>% 
    group_by(strata, community, trait) %>% 
    summarise(nGWM = weighted.mean(traitvalue, Abundance, na.rm = T)) %>% 
    mutate(NullModel = n)) %>% 
  bind_rows()
# nGWM plot
ggplot(nGWM, aes(nGWM, col = community), fill = NA) +
  geom_density(alpha = .5) +
  facet_wrap(~ strata + trait, scales = "free") +
  scale_color_discrete(guide = "none")
# SES
SES <- GWM %>% 
  left_join(group_by(nGWM, strata, community, trait) %>% 
              summarise(nGWMmean = mean(nGWM, na.rm = T),
                        nGWMsd = sd(nGWM, na.rm = T))) %>% 
  mutate(SES = (GWM - nGWMmean)/nGWMsd) %>% 
  dplyr::select(-GWM, -nGWMmean, -nGWMsd) %>% 
  left_join(R)
# SES plot
reshape2::melt(SES, id.vars = c("strata", "community", "trait", "SES")) %>% 
  filter(variable != "BA") %>% 
  ggplot(aes(value, SES, col = strata)) +
  facet_grid(trait ~ variable, scales = "free") +
  geom_point() +
  geom_hline(yintercept = c(-2, 0, 2), linetype = "dashed") +
  geom_smooth(method = "lm") +
  theme(legend.position = "bottom", axis.title = element_blank())
```

@@

```{r pca,  fig.cap='Principal Component Analysis (PCA) of species mean trait values (A) and density distribution of species per guild (C canopy, U understory) along the first two PCA axes (B). Subplot A shows PCA biplot in plan 1-2, whereas boxplots in B show guild differentiation on the first and second PCA axes, with the associated Wilcoxon test. See Table  \\@ref(tab:traits) for trait abbreviations.', eval=F}
pca.plot <- autoplot(PFT.pca, 
         data = Q %>% 
           left_join(L %>% 
                       dplyr::select(-com) %>% 
                       summarise_all(sum) %>% 
                       t() %>% 
                       data.frame() %>% 
                       rename(abundance =  ".") %>% 
                       mutate(SpCode = row.names(.))) %>% 
           filter(!is.na(WES)), 
         colour = "strata",
         size = "abundance",
         loadings.label.size = 6,
         loadings.label.colour = 'black',
         loadings = T, loadings.label = T, loadings.colour = 'black',
         loadings.label.vjust = 1.2) +
  coord_equal() +
  geom_hline(aes(yintercept = 0), col = 'green', linetype = "dotted") +
  geom_vline(aes(xintercept = 0), col = 'purple', linetype = "dotted") +
  xlab('Axe 1 − 33.66 %') + ylab('Axe 2 − 29.07 %') +
  theme(legend.position = c(1, 1), 
        legend.justification = c(1, 1)) +
  scale_color_discrete("Guild")
spectrum.boxplots <- pca.plot$data %>%
  select(Comp.1, Comp.2, strata) %>%
  rename(`Axe 1` = Comp.1, `Axe 2` = Comp.2) %>%
  melt(id.vars = "strata", variable.name = "spectrum") %>%
  ggplot(aes(strata, value, fill = strata)) +
  ggpubr::stat_compare_means(aes(label = ..p.signif..)) +
  geom_boxplot() +
  facet_wrap(~spectrum, nrow = 2) +
  scale_fill_discrete("Guild")
cowplot::plot_grid(pca.plot, spectrum.boxplots, rel_widths = c(2,1))
```

@@

```{r GWMboxplot, fig.cap="Guild weighted mean (GWM) of specific leaf area (SLA), wood density (WD), leaf dry matter content (LDMC), leaf area (LA) and leaf thickness (LT) for canopy (C) and understory (U) guilds. Boxplots include the median value, the interquartile range (horizontal edges of the box) and an estimated 95% confidence interval (vertical lines). Asterisks represent significant differences between the two guilds (*** p-value<0.001).", fig.height=4, eval=F}
L %>%
  melt(id.var = "com",
       variable.name = "SpCode",
       value.name = "abundance") %>% 
  dcast(SpCode ~ com, value.var = "abundance") %>% 
  mutate(SpCode = as.character(SpCode)) %>% 
  inner_join(Q, by = "SpCode") %>%
  select(-SpCode) %>%
  melt(id.vars = c('strata', 'LES', 'SLA', 'WES', 'WD', 'LDMC', 'LA', 'LT'),
       variable.name = 'community', value.name = 'abundance') %>%
  group_by(strata, community) %>%
  summarise_each(funs(weighted.mean(., abundance, na.rm = T)), -abundance) %>% 
  select(-community, -LES, -WES) %>%
  # select(-community, -SLA, -WD, -LDMC, -LA, -LT) %>% 
  # select(-community) %>% 
  melt(id.vars = "strata", 
       variable.name = "trait",
       value.name = "GWM") %>% 
  ggplot(aes(x = strata, y = GWM, fill = strata)) +
  geom_boxplot() +
  facet_wrap(~trait, scales = "free", nrow = 1) +
  # facet_wrap(~trait, scales = "free", nrow = 2) +
  xlab("Guild") + scale_fill_discrete(guide = "none") +
  ggpubr::stat_compare_means(aes(label = paste(substr(..p.signif.., 1,3))))
  # ggsave(file = "boxplot.png",
  #        path = './data/',
  #        width = 200, height = 150, unit = 'mm',
  #        dpi = 300)
```
